---
title: reactable Visualization of Allocation
author: Kent Russell
date: '2023-12-10'
slug: reactable-allocation
categories: []
tags: []
---

In my real job as a portfolio manager, asset allocation is very important, so we monitor on a routine basis.  For the weekly update, I used [`reactable`](https://glin.github.io/reactable) with a little custom `d3` JavaScript and SVG to visualize the current allocation compared to investment policy ranges and targets.  Below is a quick example with entirely made up (but I think realistic) numbers.


```{r,warning=FALSE,echo=FALSE}
library(htmltools)
library(d3r)
library(reactable)

# example portfolio allocation data frame
#   in general this will likely come from another source
alloc <- data.frame(
  asset = c("Fixed Income","Equity","Real Estate", "Cash"),
  val = c(1400000000, 750000000, 500000000, 100000000),
  pct = rep(NA,4), # will calculate later
  target_pct = c(0.50,0.25,0.20,0.05),
  target_val = rep(NA,4), # will calculate later
  diff = rep(NA,4), # will calculate later
  min = c(0.40,0.20,0.15,0.02),
  max = c(0.60,0.30,0.25,0.10)
)

# calculate actual percent
alloc$pct <- alloc$val / sum(alloc$val)

# calculate target value
alloc$target_val <- sum(alloc$val) * alloc$target_pct

# calculate actual to target difference
alloc$diff <- alloc$val - alloc$target_val

rt <- reactable(
  alloc,
  pagination = FALSE,
  sortable = FALSE,
  compact = TRUE,
  style = list(fontFamily = "sans-serif"),
  # turn off borders here
  borderless = TRUE,
  rowStyle = list(
    fontSize = "1.25rem",
    alignItems = "center",
    # add back row borders here
    borderBottom = "1px solid lightgray"
  ),
  defaultColDef = colDef(
    headerStyle = "align-self: flex-end; font-weight: normal; text-align:center;"
  ),
  columns = list(
    asset = colDef(
      header = JS("function(cellInfo){return `<div></div>`}"),
      headerStyle = "text-transform:uppercase; align-self:flex-end; font-weight:normal;",
      html = TRUE,
      cell = JS("
function(cellInfo) {
  return `
    <div>
        ${cellInfo.value}
    </div>
  `
}
      ")
    ),
    val = colDef(
      header = JS("function(cellInfo){return `<div>$</div>`}"),
      cell = JS("
function(cellInfo) {
  if(cellInfo.index === 0) {
    return `<div>${'$ ' + d3.format(',.0f')(cellInfo.value)}</div>`
  } else {
    return `<div>${d3.format(',.0f')(cellInfo.value)}</div>`
  }
}
      "),
      html = TRUE
    ),
    pct = colDef(
      header = JS("function(cellInfo){return `<div>%</div>`}"),
      cell = JS("
function(cellInfo) {
  if(cellInfo.row.asset === '') return null
  const height = 50
  const width = 400
  const pady = 12
  const padx = 20
  return `
<svg height = ${height} width = ${width}>
  <rect
    x='${scalex(cellInfo.row.min)}'
    y='${pady}'
    width='${scalex(cellInfo.row.max)-scalex(cellInfo.row.min)}'
    height='${height-(2*pady)}'
    fill='#ccc'
  />
  <line
    x1='${scalex(cellInfo.row.target_pct)}'
    y1='${pady}'
    x2='${scalex(cellInfo.row.target_pct)}'
    y2='${height-pady}'
    stroke='white'
    stroke-width='2'
  />
  <line
    x1='${scalex(cellInfo.row.pct)}'
    y1='${pady}'
    x2='${scalex(cellInfo.row.pct)}'
    y2='${height-pady}'
    stroke='#4fb1ff'
    stroke-width='2'
  />
  <text
    x='${scalex(cellInfo.row.min)-2}'
    y='${height/2 + 4}'
    text-anchor='end'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.min)}
  </text>
  <text
    x='${scalex(cellInfo.row.max)+2}'
    y='${height/2 + 4}'
    text-anchor='start'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.max)}
  </text>
  <text
    x='${scalex(cellInfo.row.target_pct)}'
    y='${height - pady/2 + 4}'
    text-anchor='middle'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.target_pct)}
  </text>
  <text
    x='${scalex(cellInfo.row.pct)}'
    y='${pady - 4}'
    text-anchor='middle'
    fill='#4fb1ff'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.pct)}
  </text>
  <text
    x='${scalex(1) + 20}'
    y='${height/2 + 20}'
    text-anchor='end'
    fill='${cellInfo.row.pct > cellInfo.row.target_pct ? 'black' : 'red'}'
    style='font-size:20px;'
  >
    ${
      cellInfo.row.pct > cellInfo.row.target_pct ?
        '+' + d3.format('.1%')(cellInfo.row.pct - cellInfo.row.target_pct) :
        d3.format('.1%')(cellInfo.row.pct - cellInfo.row.target_pct)
    }
  </text>
</svg>
  `
}
      "),
      html = TRUE,
      width = 400
    ),
    target_pct = colDef(show = FALSE),
    target_val = colDef(show = FALSE),
    diff = colDef(show = FALSE),
    min = colDef(show = FALSE),
    max = colDef(show = FALSE)
  )
)

tagList(
  d3r::d3_dep_v7(),
  tags$script(HTML("const scalex = d3.scaleLinear().range([40,360])")),
  tags$div(
    style = "width: 800px; background: white;",
    rt
  )
)
```


### Code


```{r,eval=FALSE}
library(htmltools)
library(d3r)
library(reactable)

# example portfolio allocation data frame
#   in general this will likely come from another source
alloc <- data.frame(
  asset = c("Fixed Income","Equity","Real Estate", "Cash"),
  val = c(1400000000, 750000000, 500000000, 100000000),
  pct = rep(NA,4), # will calculate later
  target_pct = c(0.50,0.25,0.20,0.05),
  target_val = rep(NA,4), # will calculate later
  diff = rep(NA,4), # will calculate later
  min = c(0.40,0.20,0.15,0.02),
  max = c(0.60,0.30,0.25,0.10)
)

# calculate actual percent
alloc$pct <- alloc$val / sum(alloc$val)

# calculate target value
alloc$target_val <- sum(alloc$val) * alloc$target_pct

# calculate actual to target difference
alloc$diff <- alloc$val - alloc$target_val

rt <- reactable(
  alloc,
  pagination = FALSE,
  sortable = FALSE,
  compact = TRUE,
  style = list(fontFamily = "sans-serif"),
  # turn off borders here
  borderless = TRUE,
  rowStyle = list(
    fontSize = "1.25rem",
    alignItems = "center",
    # add back row borders here
    borderBottom = "1px solid lightgray"
  ),
  defaultColDef = colDef(
    headerStyle = "align-self: flex-end; font-weight: normal; text-align:center;"
  ),
  columns = list(
    asset = colDef(
      header = JS("function(cellInfo){return `<div></div>`}"),
      headerStyle = "text-transform:uppercase; align-self:flex-end; font-weight:normal;",
      html = TRUE,
      cell = JS("
function(cellInfo) {
  return `
    <div>
        ${cellInfo.value}
    </div>
  `
}
      ")
    ),
    val = colDef(
      header = JS("function(cellInfo){return `<div>$</div>`}"),
      cell = JS("
function(cellInfo) {
  if(cellInfo.index === 0) {
    return `<div>${'$ ' + d3.format(',.0f')(cellInfo.value)}</div>`
  } else {
    return `<div>${d3.format(',.0f')(cellInfo.value)}</div>`
  }
}
      "),
      html = TRUE
    ),
    pct = colDef(
      header = JS("function(cellInfo){return `<div>%</div>`}"),
      cell = JS("
function(cellInfo) {
  if(cellInfo.row.asset === '') return null
  const height = 50
  const width = 400
  const pady = 12
  const padx = 20
  return `
<svg height = ${height} width = ${width}>
  <rect
    x='${scalex(cellInfo.row.min)}'
    y='${pady}'
    width='${scalex(cellInfo.row.max)-scalex(cellInfo.row.min)}'
    height='${height-(2*pady)}'
    fill='#ccc'
  />
  <line
    x1='${scalex(cellInfo.row.target_pct)}'
    y1='${pady}'
    x2='${scalex(cellInfo.row.target_pct)}'
    y2='${height-pady}'
    stroke='white'
    stroke-width='2'
  />
  <line
    x1='${scalex(cellInfo.row.pct)}'
    y1='${pady}'
    x2='${scalex(cellInfo.row.pct)}'
    y2='${height-pady}'
    stroke='#4fb1ff'
    stroke-width='2'
  />
  <text
    x='${scalex(cellInfo.row.min)-2}'
    y='${height/2 + 4}'
    text-anchor='end'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.min)}
  </text>
  <text
    x='${scalex(cellInfo.row.max)+2}'
    y='${height/2 + 4}'
    text-anchor='start'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.max)}
  </text>
  <text
    x='${scalex(cellInfo.row.target_pct)}'
    y='${height - pady/2 + 4}'
    text-anchor='middle'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.target_pct)}
  </text>
  <text
    x='${scalex(cellInfo.row.pct)}'
    y='${pady - 4}'
    text-anchor='middle'
    fill='#4fb1ff'
    style='font-size:11px;'
  >
    ${d3.format('.0%')(cellInfo.row.pct)}
  </text>
  <text
    x='${scalex(1) + 20}'
    y='${height/2 + 20}'
    text-anchor='end'
    fill='${cellInfo.row.pct > cellInfo.row.target_pct ? 'black' : 'red'}'
    style='font-size:20px;'
  >
    ${
      cellInfo.row.pct > cellInfo.row.target_pct ?
        '+' + d3.format('.1%')(cellInfo.row.pct - cellInfo.row.target_pct) :
        d3.format('.1%')(cellInfo.row.pct - cellInfo.row.target_pct)
    }
  </text>
</svg>
  `
}
      "),
      html = TRUE,
      width = 400
    ),
    target_pct = colDef(show = FALSE),
    target_val = colDef(show = FALSE),
    diff = colDef(show = FALSE),
    min = colDef(show = FALSE),
    max = colDef(show = FALSE)
  )
)

browsable(tagList(
  d3r::d3_dep_v7(),
  tags$script(HTML("const scalex = d3.scaleLinear().range([40,360])")),
  tags$div(
    style = "width: 800px; background: white;",
    rt
  )
))
